Continuous time

### 2023-09-21 continuous-time.qmd
edit text
and some refactoring

path <- "models/lmm-ns4.Rds"
if (file.exists(path)) {
  fit <- read_rds(path)
} else {
  fit <- lmer(
    mood100 ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
    data = dat
  )
  write_rds(fit, path)
}

@@

```{r}
#| label: fig-plot-curve
#| fig-cap: Estimated mood during a play session

hours <- seq(0, 3, length.out = 100)
emm1 <- emmeans(
  fit,
  ~hours,
  at = list(hours = hours),
  lmer.df = "asymptotic"
)
p_curve <- as.data.frame(emm1) |>
  ggplot(aes(hours, emmean)) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL
    ),
    alpha = 0.25
  ) +
  labs(
    x = "Session duration (hours)",
    y = "Mood (0 - 100)"
  )
p_curve
```

@@

```{r}
#| label: fig-plot-change
#| fig-cap: Estimated difference in mood over a session compared to the start of a session

d_change <- emmeans::contrast(
  emm1,
  method = "trt.vs.ctrl",
  ref = "hours0"
) |>
  confint() |>
  as.data.frame() |>
  mutate(
    hours = hours[-1]
  )

@@

```{r}
#| label: fig-intercept-sds
#| fig-cap: Estimated curves for participants with an intercept equal to 0, -1SD, and -2SD.
get_trend <- function(
    fit,
    data,
    sd = 0,
    intercept = TRUE) {
  re <- as.data.frame(VarCorr(fit))
  pid_int <- re[2, "sdcor"]
  pid_slope0 <- re[3, "sdcor"]
  pid_slope1 <- re[4, "sdcor"]
  pid_slope2 <- re[5, "sdcor"]
  pid_slope3 <- re[6, "sdcor"]
  pid_cor0 <- re[7, "sdcor"]
  pid_cor1 <- re[8, "sdcor"]
  pid_cor2 <- re[9, "sdcor"]
  pid_cor3 <- re[10, "sdcor"]
  re_int <- sd * pid_int
  re_s0 <- sd * pid_cor0 * pid_slope0
  re_s1 <- sd * pid_cor1 * pid_slope1
  re_s2 <- sd * pid_cor2 * pid_slope2
  re_s3 <- sd * pid_cor3 * pid_slope3
  hours <- seq(0, 3, length.out = 100)
  x <- ns4(
    hours
  )
  b0 <- fixef(fit)[1] + re_int
  b <- fixef(fit)[-1]
  y <- x[, 1] * (b[1] + re_s0) + x[, 2] * (b[2] + re_s1) + x[, 3] * (b[3] + re_s2) + x[, 4] * (b[4] + re_s3)
  if (intercept) y <- y + b0
  data.frame(
    x = hours,
    y = c(y),
    re_int
  )
}
color_scale <- list(
  scale_color_manual(
    "Intercept",
    values = c(
      "1 SD" = "#4daf4a",  
      "-1 SD" = "#e41a1c",
      "-2 SD" = "#377eb8",
      "0" = "#000000"
    )
  )
)

p_int_sd <- rbind(
get_trend(
    fit,
    data = dat,
    sd = 1
  ),
  get_trend(fit, data = dat),
  get_trend(
    fit,
    data = dat,
    sd = -1
  ),
  get_trend(
    fit,
    data = dat,
    sd = -2
  )
) |>
  mutate(
    intercept = factor(
      rep(
        c("1 SD", "0", "-1 SD", "-2 SD"),
        each = 100
      )
    )
  ) |>
  ggplot(
    aes(
      x,
      y,
      color = intercept,
      group = intercept
    )
  ) +
  geom_line() +
  labs(
    x = "Session duration (hours)",
    y = "Mood (0 - 100)"
  ) +
  color_scale

@@

```{r}
#| label: fig-no-intercept
#| fig-cap: Estimated change curves for participants with an intercept equal to 0, -1SD, and -2SD.
p_int_sd_change <- rbind(
  get_trend(fit, data = dat, intercept = FALSE),
  get_trend(
    fit,
    data = dat,
    sd = -1,
    intercept = FALSE
  ),
  get_trend(
    fit,
    data = dat,
    sd = -2,
    intercept = FALSE
  )
) |>
  mutate(
    intercept = factor(
      rep(
        c("0", "-1 SD", "-2 SD"),
        each = 100
      )
    )
  ) |>
  ggplot(
    aes(
      x,
      y,
      color = intercept,
      group = intercept
    )
  ) +
  geom_line() +
  geom_ribbon(
    data = d_change,
    aes(
      x = hours,
      y = NULL,
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      group = "0",
      color = NULL
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  labs(
    x = "Session duration (hours)",
    y = "Change in Mood (0 - 100)"
  ) +
  color_scale
p_int_sd_change
```

### Combined figured
Combined figure for manuscript(?).

```{r}
#| label: prod-figure
#| fig-width: 8
#| fig-height: 5
p_comb <- (p_curve) /
  (
    p_int_sd +
      p_int_sd_chage +
      plot_layout(guides = "collect") &
      theme(legend.position = "bottom")
  ) +
  plot_annotation(tag_levels = "A")

@@

@@

```{r}
#| label: lmm-reduced

fit2 <- lmer(
  mood ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
  data = fit_b$data
)
```

@@

```{r}
#| label: fig-cens-compare
#| fig-cap: Estimated trends with and without censoring.
d_brms <-
  emmeans(
    fit_b,
    ~hours,
    at = list(hours = seq(0, 3, length.out = 100))
  ) |>
  as.data.frame() |>
  mutate(model = "cens") |>
  select(
    hours,
    emmean,
    model,
    lower = "lower.HPD",
    upper = "upper.HPD"
  )
# SHIFT <- 0.017
SHIFT <- 0
d_lmm <-
  emmeans(
    fit2,
    ~hours,
    at = list(hours = seq(0, 3, length.out = 100))
  ) |>
  as.data.frame() |>
  select(
    hours,
    emmean,
    lower = "asymp.LCL",
    upper = "asymp.UCL"
  ) |>
  mutate(
    model = "lmm",
    emmean = emmean + SHIFT,
    upper = upper + SHIFT,
    lower = lower + SHIFT
  )
rbind(d_brms, d_lmm) |>
  ggplot(
    aes(
      hours,
      emmean,
      group = model,
      color = model
    )
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = lower,
      ymax = upper,
      fill = model,
      color = NULL
    ),
    alpha = 0.25
  ) +
  labs(x = "Session duration (hours)")
```

@@

get_trend_bayes <- function(sd) {
  re <- VarCorr(fit_b)
  ps <- as_draws_df(
    fit_b,
    c(
      "b_Intercept",
      "b_ns4hours1",
      "b_ns4hours2",
      "b_ns4hours3",
      "b_ns4hours4",
      "sd_pid__Intercept",
      "sd_pid__ns4hours1",
      "sd_pid__ns4hours2",
      "sd_pid__ns4hours3",
      "sd_pid__ns4hours4",
      "cor_pid__Intercept__ns4hours1",
      "cor_pid__Intercept__ns4hours2",
      "cor_pid__Intercept__ns4hours3",
      "cor_pid__Intercept__ns4hours4"
    )
  ) |>
    rename(
      pid_int = sd_pid__Intercept,
      pid_slope0 = sd_pid__ns4hours1,
      pid_slope1 = sd_pid__ns4hours2,
      pid_slope2 = sd_pid__ns4hours3,
      pid_slope3 = sd_pid__ns4hours4,
      pid_cor0 = cor_pid__Intercept__ns4hours1,
      pid_cor1 = cor_pid__Intercept__ns4hours2,
      pid_cor2 = cor_pid__Intercept__ns4hours3,
      pid_cor3 = cor_pid__Intercept__ns4hours4,
      b1 = b_ns4hours1,
      b2 = b_ns4hours2,
      b3 = b_ns4hours3,
      b4 = b_ns4hours4,
    ) |>
    mutate(
      re_int = sd * pid_int,
      b0 = b_Intercept + re_int,
      re_s0 = sd * pid_cor0 * pid_slope0,
      re_s1 = sd * pid_cor1 * pid_slope1,
      re_s2 = sd * pid_cor2 * pid_slope2,
      re_s3 = sd * pid_cor3 * pid_slope3
    )
  lapply(hours, function(hour) {
    x <- ns4(
      hour
    )
    ps |>
      mutate(
        y = b0 + x[, 1] * (b1 + re_s0) + x[, 2] * (b2 + re_s1) + x[, 3] * (b3 + re_s2) + x[, 4] * (b4 + re_s3)
      ) |>
      select(y) |>
      summarize(
        estimate = median(y),
        lwr = quantile(y, 0.025),
        upr = quantile(y, 0.975)
      ) |>
      mutate(
        hours = hour,
        re_int = mean(ps$re_int)
      )
  }) |>
    bind_rows() |>
    mutate(sd = sd)
}
res <- bind_rows(
  get_trend_bayes(0),
  get_trend_bayes(-1),
  get_trend_bayes(-2),
)
res <- res |>
  mutate(
    intercept = factor(
      sd,
      levels = c(0, -1, -2),
      labels = c("0", "-1 SD", "-2 SD")
    )
  )
# lmm for comparison
dat2 <- fit_b$data
res_lmm <- rbind(
  get_trend(fit2, data = dat2),
  get_trend(
    fit2,
    data = dat2,
    sd = -1
  ),
  get_trend(
    fit2,
    data = dat2,
    sd = -2
  )
) |>
  mutate(
    intercept = factor(
      rep(
        c("0", "-1 SD", "-2 SD"),
        each = 100
      )
    )
  ) |>
  select(
    estimate = y,
    hours = x,
    intercept
  )

@@

```{r}
#| label: fig-bayes-int-sd
#| fig-cap: Bayesian censored normal model vs LMM (dashed) fit to subset of data

ggplot(
  res,
  aes(
    hours, estimate,
    group = intercept,
    color = intercept
  )
) +
  geom_line() +
  geom_line(
    data = res_lmm,
    linetype = "dashed"
  ) +
  geom_ribbon(
    aes(
      x = hours,
      y = NULL,
      ymin = lwr,
      ymax = upr,
      color = NULL
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) + 
  color_scale
```

### Pretest x time continuous interaction
To check the robustness of the impact of baseline wellbeing, we also fit a separate interaction model to the subset of participants that began their session by answering a mood question. This allows us to explicitly model the interaction between the pretest values and the within-session change, without having to infer this from the multivariate normal distribution of the random effects.

Plot the interaction at different pretest values. As shown in @fig-lmm-pre-interaction, this interaction model shows a pattern similar to @fig-intercept-sds where those who report lower mood at the beginning of a session also improve more. 
```{r}
#| label: fig-lmm-pre-interaction
#| fig-cap: Interaction between pretest mood and change, shown at -2 SD, -1, SD, the mean, and +1 SD.
pre_sum <- dat2 |> 
    summarise(
        pre_mean = median(pre), .by = pid
    ) |> 
    summarise(
        mean = mean(pre_mean), sd = sd(pre_mean)
    )
emm <- emmeans(
    fit3,
    ~hours + pre,
    at = list(
        hours = seq(0, 3, length.out = 100), 
        pre = c(
            pre_sum$mean - 2 * pre_sum$sd, 
            pre_sum$mean - pre_sum$sd, 
            pre_sum$mean, 
            pre_sum$mean + pre_sum$sd)
            ),
    lmer.df = "asymp"
)
p <- as.data.frame(emm) |> 
    mutate(
        pre = factor(round(pre))
    ) |> 
    ggplot(
        aes(hours, emmean, group = pre, color = pre)
    ) + 
    geom_line() +
    geom_ribbon(
        aes(ymin = asymp.LCL, ymax = asymp.UCL, fill = pre, color = NULL),
        alpha = 0.25
        ) 
p
ggsave("figures/fig-interaction-model.png", p, width = 7, height = 5)
```

(Ignore this change model --- not sure if it's worth recreating it).
```{r}
#| eval: false
#| label: fig-lmm-pre-interaction-change
emm <- emmeans(
    fit3,
    ~hours + pre,
    at = list(
        hours = seq(0, 3, length.out = 100), 
        pre = c(
            pre_sum$mean)
            ),
    lmer.df = "asymp"
)
d_change <- emmeans::contrast(
  emm,
  method = "trt.vs.ctrl",
  ref = "hours0 pre72.2106770833333"
) |>
  confint() |>
  as.data.frame() |>
  mutate(
    hours = hours[-1]
  )

p_change <- d_change |>
  ggplot(aes(hours, estimate)) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL
    ),
    alpha = 0.25
  ) +
  labs(x = "Session duration (hours)")
p_change
```

For completeness we'll also fit the censored normal model with 1000 participants.

@@

We then compare the censored normal model with the LMM and recreate @fig-lmm-pre-interaction. As shown in @fig-cens-pre-interaction, there's almost no difference between the models. The LMMs CIs are slightly narrower near the intercept.
```{r}
#| label: fig-cens-pre-interaction
#| fig-cap: Censored normal vs LMM with a continuous by continuous interaction between within-session change and pretest mood.
get_emm_pre <- function(fit) {
  emmeans(
    fit,
    ~ hours + pre,
    # data = dat2,
    at = list(
      hours = seq(0, 3, length.out = 100),
      pre = c(
        pre_sum$mean - 2 * pre_sum$sd,
        pre_sum$mean - pre_sum$sd,
        pre_sum$mean,
        pre_sum$mean + pre_sum$sd
      )
    ),
    lmer.df = "asymp"
  )
}
pre_sum <- dat2 |>
  summarise(
    pre_mean = median(pre), .by = pid
  ) |>
  summarise(
    mean = mean(pre_mean), sd = sd(pre_mean)
  )
emm_cens <- get_emm_pre(fit3_b) |>
  as.data.frame() |>
  mutate(
    model = "cens"
  ) |>
  rename(
    lwr = lower.HPD,
    upr = upper.HPD
  )
emm_lmm <- get_emm_pre(fit3_small) |>
  as.data.frame() |>
  mutate(
    model = "LMM"
  ) |>
  rename(
    lwr = asymp.LCL,
    upr = asymp.UCL
  )

p <- bind_rows(emm_cens, emm_lmm) |>
  mutate(
    pre = factor(
      pre,
      levels = sort(unique(pre)),
      labels = c("-2 SD", "-1 SD", "Mean", "+1 SD")
    )
  ) |>
  {
    \(d)
    ggplot(
      data = d,
      aes(
        hours,
        emmean,
        group = interaction(pre, model),
        color = pre,
        linetype = model
      )
    ) +
      geom_line() +
      geom_ribbon(
        data = filter(d, model == "cens"),
        aes(
          ymin = lwr,
          ymax = upr,
          fill = pre,
          color = pre,
        ),
        color = NA,
        alpha = 0.25
      ) +
      geom_ribbon(
        data = filter(d, model == "LMM"),
        aes(
          ymin = lwr,
          ymax = upr,
          fill = pre,
          color = pre,
        ),
        fill = NA,
        alpha = 0.15
      ) +
      scale_color_brewer(palette = "Set1") +
      scale_fill_brewer(palette = "Set1")
  }()
p
```

>>>>

path <- "models/lmm-ns4.Rds"
if (file.exists(path)) {
  fit <- read_rds(path)
} else {
  fit <- lmer(
    mood100 ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
    data = dat
  )
  write_rds(fit, path)
}
```

This model included `r number2(ngrps(fit)["pid"])` participants, with a total of `r number2(ngrps(fit)["ps"])` sessions and `r number2(nobs(fit))` observations.

@@

```{r}
#| label: fig-plot-curve
#| fig-cap: Estimated mood during a play session
#| fig-height: 3
#| fig-width: 8
hours <- seq(0, 3, length.out = 100)
emm1 <- emmeans(
  fit,
  ~hours,
  at = list(hours = hours),
  lmer.df = "asymptotic"
)
p_curve <- as.data.frame(emm1) |>
  ggplot(aes(hours, emmean)) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL
    ),
    alpha = 0.25
  ) +
  labs(
    x = "Session duration (hours)",
    y = "Mood (0 - 100)"
  )
p_curve
```

@@

```{r}
#| label: fig-plot-change
#| fig-cap: Estimated difference in mood over a session compared to the start of a session
#| fig-height: 3
#| fig-width: 8
d_change <- emmeans::contrast(
  emm1,
  method = "trt.vs.ctrl",
  ref = "hours0"
) |>
  confint() |>
  as.data.frame() |>
  mutate(
    hours = hours[-1]
  )

@@

```{r}
#| label: fig-intercept-sds
#| fig-cap: Estimated curves for participants with an intercept equal to 0, -1SD, and -2SD.
#| fig-height: 3
#| fig-width: 8
get_trend <- function(
    fit,
    data,
    sd = 0,
    intercept = TRUE) {
  re <- as.data.frame(VarCorr(fit))
  pid_int <- re[2, "sdcor"]
  pid_slope0 <- re[3, "sdcor"]
  pid_slope1 <- re[4, "sdcor"]
  pid_slope2 <- re[5, "sdcor"]
  pid_slope3 <- re[6, "sdcor"]
  pid_cor0 <- re[7, "sdcor"]
  pid_cor1 <- re[8, "sdcor"]
  pid_cor2 <- re[9, "sdcor"]
  pid_cor3 <- re[10, "sdcor"]
  re_int <- sd * pid_int
  re_s0 <- sd * pid_cor0 * pid_slope0
  re_s1 <- sd * pid_cor1 * pid_slope1
  re_s2 <- sd * pid_cor2 * pid_slope2
  re_s3 <- sd * pid_cor3 * pid_slope3
  hours <- seq(0, 3, length.out = 100)
  x <- ns4(
    hours
  )
  b0 <- fixef(fit)[1] + re_int
  b <- fixef(fit)[-1]
  y <- x[, 1] * (b[1] + re_s0) + x[, 2] * (b[2] + re_s1) + x[, 3] * (b[3] + re_s2) + x[, 4] * (b[4] + re_s3)
  if (intercept) y <- y + b0
  data.frame(
    x = hours,
    y = c(y),
    re_int
  )
}

pre_labels <- c("-2 SD", "-1 SD", "Mean", "1 SD")
pre_values <- c(-2, -1, 0, 1)
colors <- c(
  "1 SD" = "#4daf4a",
  "-1 SD" = "#e41a1c",
  "-2 SD" = "#377eb8",
  "Mean" = "#000000"
)
color_scale <- list(
  scale_color_manual(
    "Pre",
    values = colors
  ),
    scale_fill_manual(
    "Pre",
    values = colors
  )
)

p_int_sd <- map(
  pre_values,
  \(sd) get_trend(fit, dat, sd = sd)
) |>
  bind_rows() |>
  mutate(
    intercept = factor(
      rep(
        pre_labels,
        each = 100
      ),
      levels = pre_labels
    )
  ) |>
  ggplot(
    aes(
      x,
      y,
      color = intercept,
      group = intercept
    )
  ) +
  geom_line() +
  labs(
    x = "Session duration (hours)",
    y = "Mood (0 - 100)"
  ) +
  color_scale

@@

```{r}
#| label: fig-no-intercept
#| fig-cap: Estimated change curves for participants with an intercept equal to 0, -1SD, and -2SD.
#| fig-height: 3
#| fig-width: 8
p_int_sd_change <- map(
  pre_values,
  \(sd) get_trend(fit, dat, sd = sd, intercept = FALSE)
) |>
  bind_rows() |>
  mutate(
    intercept = factor(
      rep(
        pre_labels,
        each = 100
      ),
      levels = pre_labels
    )
  ) |>
  ggplot(
    aes(
      x,
      y,
      color = intercept,
      group = intercept
    )
  ) +
  geom_line() +
  geom_ribbon(
    data = d_change,
    aes(
      x = hours,
      y = NULL,
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      group = "0",
      color = NULL
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  labs(
    x = "Session duration (hours)",
    y = "Change in Mood (0 - 100)"
  ) +
  color_scale
p_int_sd_change
```

```{r}
#| label: fig-session-change
#| echo: false
#| fig-width: 8
#| fig-height: 5
p_comb <- (p_curve) /
  (
    p_int_sd +
      p_int_sd_change +
      plot_layout(guides = "collect") &
      theme(legend.position = "bottom")
  ) +
  plot_annotation(tag_levels = "A")

@@

### Pretest x time continuous interaction
Modeling the relationship between baseline wellbeing and change using random effects is limited. Here we'll fit a separate interaction model to the subset of participants that began their session by answering a mood question. This allows us to explicitly model the interaction between the pretest values and the within-session change, without having to infer this from the multivariate normal distribution of the random effects. This also allows us to model this as a non-linear relationship.

This model included `r number2(ngrps(fit3)["pid"])` participants, with a total of `r number2(ngrps(fit3)["ps"])` sessions and `r number2(nobs(fit3))` observations.

Plot the interaction at different pretest values. As shown in @fig-lmm-pre-interaction, this interaction model shows a pattern similar to @fig-intercept-sds where those who report lower mood at the beginning of a session also improve more. 
```{r}
#| label: fig-lmm-pre-interaction
#| fig-cap: Interaction between pretest mood and change, shown at -2 SD, -1, SD, the mean, and +1 SD.
#| fig-height: 3
#| fig-width: 8
pre_sum <- dat2 |>
  summarise(
    pre_mean = median(pre), .by = pid
  ) |>
  summarise(
    mean = mean(pre_mean), sd = sd(pre_mean)
  )
emm <- emmeans(
  fit3,
  ~ hours + pre,
  at = list(
    hours = seq(0, 3, length.out = 100),
    pre = pre_sum$mean + pre_values * pre_sum$sd
  ),
  lmer.df = "asymp"
)
p_pre_x <- as.data.frame(emm) |>
  mutate(
    pre = factor(
      round(pre), 
      labels = pre_labels
      )
  ) |>
  ggplot(
    aes(
      hours,
      emmean,
      group = pre,
      color = pre
    )
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      fill = pre,
      color = NULL
    ),
    alpha = 0.25,
    show.legend = FALSE
  ) + color_scale
p_pre_x
ggsave(
  "figures/fig-interaction-model.png", 
  p_pre_x, 
  width = 7, 
  height = 5
)
```

Let's also calculate the within-session change for the interaction model.
```{r}
#| label: fig-lmm-pre-interaction-change
#| fig-cap: Within-session change from the interaction model
#| fig-height: 2.5
#| fig-width: 8
get_change_by_pre <- function(pre, label, fit, hours) {
  emm <- emmeans(
    fit,
    ~ hours + pre,
    at = list(
      hours = hours,
      pre = pre
    ),
    lmer.df = "asymp"
  )
  emmeans::contrast(
    emm,
    method = "trt.vs.ctrl",
    ref = paste0("hours0 pre", pre)
  ) |>
    confint() |>
    as.data.frame() |>
    mutate(
      hours = hours[-1],
      pre = pre,
      label = label
    )
}

d_x_change <- data.frame(
  pre = round(pre_sum$mean + pre_values * pre_sum$sd),
  label = pre_labels
) |>
  pmap(\(pre, label) get_change_by_pre(
    pre,
    label,
    fit3,
    hours = seq(0, 3, length.out = 100)
  )) |>
  bind_rows()

p_pre_x_change <- d_x_change |>
  mutate(
    label = factor(
      label,
      levels = pre_labels
    )
  ) |>
  ggplot(
    aes(
      hours,
      estimate,
      group = label,
      color = label,
      fill = label
    )
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      color = NULL
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  geom_hline(
    yintercept = 0,
    linetype = "dotted"
  ) +
  labs(x = "Session duration (hours)") +
  color_scale +
  facet_wrap(~label, ncol = 4)
p_pre_x_change
```

### Manuscript figures
Create figures for the manuscript.
```{r}
#| fig-height: 4
#| fig-width: 8
p_tmp <- (
  (
    p_pre_x + 
      facet_wrap(~pre, ncol = 4) +
      labs(y = "Mood", x = "Hours")
    ) / 
    (
      p_pre_x_change +
        labs(y = "Change in mood", x = "Hours")
      )
    ) +
   plot_annotation(tag_levels = "A") +
  plot_layout(
    heights = c(2, 2),
    widths = c(8, 8),
    guides = "collect"
  ) & theme(legend.position = "bottom") 
p_tmp
```

```{r}
#| echo: false
ggsave(
  "figures/fig-pre-x-change.png", 
  p_tmp, 
  width = 8, 
  height = 4
  )
```

And this figure.

```{r}
#| fig-height: 4
#| fig-width: 8
p_tmp <- (
  p_curve + labs(x = "Hours", y = "Mood")
  ) / (
  p_change + labs(x = "Hours", y = "Change in mood")
  ) + plot_annotation(tag_levels = "A") 
p_tmp
```

```{r}
#| echo: false
ggsave(
  "figures/fig-change.png", 
  p_tmp, 
  width = 8, 
  height = 4
  )
```

And a table.
```{r}
#| label: tbl-pre-interaction-change
#| tbl-cap: Estimated within-session change from the interaction model
pre_sds <- c(1, 0, -1, -2)
tbl_x_change <- data.frame(
  pre = round(pre_sum$mean + pre_values * pre_sum$sd),
  label = pre_labels
) |>
  pmap(\(pre, label) get_change_by_pre(
    pre, 
    label, 
    fit3, 
    hours = c(0, 10/60, 0.5, 1, 2))
    ) |>
  bind_rows() |> 
  select(
    label, 
    pre, 
    hours,
    estimate, 
    CI_lwr = asymp.LCL, 
    CI_upr = asymp.UCL
    )
tbl_x_change |> 
  kable(digits = 2)
```

@@

```{r}
#| label: lmm-reduced
fit2 <- lmer(
  mood ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
  data = fit_b$data
)
```

@@

```{r}
#| label: fig-cens-compare
#| fig-cap: Estimated trends with and without censoring.
#| fig-height: 3
#| fig-width: 8
d_brms <-
  emmeans(
    fit_b,
    ~hours,
    at = list(hours = seq(0, 3, length.out = 100))
  ) |>
  as.data.frame() |>
  mutate(model = "cens") |>
  select(
    hours,
    emmean,
    model,
    lower = "lower.HPD",
    upper = "upper.HPD"
  )
# SHIFT <- 0.017
SHIFT <- 0
d_lmm <-
  emmeans(
    fit2,
    ~hours,
    at = list(hours = seq(0, 3, length.out = 100))
  ) |>
  as.data.frame() |>
  select(
    hours,
    emmean,
    lower = "asymp.LCL",
    upper = "asymp.UCL"
  ) |>
  mutate(
    model = "lmm",
    emmean = emmean + SHIFT,
    upper = upper + SHIFT,
    lower = lower + SHIFT
  )
rbind(d_brms, d_lmm) |>
  ggplot(
    aes(
      hours,
      emmean,
      group = model,
      color = model
    )
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = lower,
      ymax = upper,
      fill = model,
      color = NULL
    ),
    alpha = 0.25
  ) +
  labs(x = "Session duration (hours)")
```

@@

get_trend_bayes <- function(sd) {
  re <- VarCorr(fit_b)
  ps <- as_draws_df(
    fit_b,
    c(
      "b_Intercept",
      "b_ns4hours1",
      "b_ns4hours2",
      "b_ns4hours3",
      "b_ns4hours4",
      "sd_pid__Intercept",
      "sd_pid__ns4hours1",
      "sd_pid__ns4hours2",
      "sd_pid__ns4hours3",
      "sd_pid__ns4hours4",
      "cor_pid__Intercept__ns4hours1",
      "cor_pid__Intercept__ns4hours2",
      "cor_pid__Intercept__ns4hours3",
      "cor_pid__Intercept__ns4hours4"
    )
  ) |>
    rename(
      pid_int = sd_pid__Intercept,
      pid_slope0 = sd_pid__ns4hours1,
      pid_slope1 = sd_pid__ns4hours2,
      pid_slope2 = sd_pid__ns4hours3,
      pid_slope3 = sd_pid__ns4hours4,
      pid_cor0 = cor_pid__Intercept__ns4hours1,
      pid_cor1 = cor_pid__Intercept__ns4hours2,
      pid_cor2 = cor_pid__Intercept__ns4hours3,
      pid_cor3 = cor_pid__Intercept__ns4hours4,
      b1 = b_ns4hours1,
      b2 = b_ns4hours2,
      b3 = b_ns4hours3,
      b4 = b_ns4hours4,
    ) |>
    mutate(
      re_int = sd * pid_int,
      b0 = b_Intercept + re_int,
      re_s0 = sd * pid_cor0 * pid_slope0,
      re_s1 = sd * pid_cor1 * pid_slope1,
      re_s2 = sd * pid_cor2 * pid_slope2,
      re_s3 = sd * pid_cor3 * pid_slope3
    )
  lapply(hours, function(hour) {
    x <- ns4(
      hour
    )
    ps |>
      mutate(
        y = b0 + x[, 1] * (b1 + re_s0) + x[, 2] * (b2 + re_s1) + x[, 3] * (b3 + re_s2) + x[, 4] * (b4 + re_s3)
      ) |>
      select(y) |>
      summarize(
        estimate = median(y),
        lwr = quantile(y, 0.025),
        upr = quantile(y, 0.975)
      ) |>
      mutate(
        hours = hour,
        re_int = mean(ps$re_int)
      )
  }) |>
    bind_rows() |>
    mutate(sd = sd)
}

res <- map(
  pre_values,
  \(sd) get_trend_bayes(sd = sd)
) |>
  bind_rows() |>
  mutate(
    intercept = factor(
      sd,
      levels = pre_values,
      labels = pre_labels
    )
  )

# lmm for comparison
dat2 <- fit_b$data
res_lmm <- map(
  pre_values,
  \(sd) get_trend(fit2, dat2, sd = sd)
) |>
  bind_rows() |>
  mutate(
    intercept = factor(
      rep(
        pre_labels,
        each = 100
      ),
      levels = pre_labels
    )
  ) |>
  select(
    estimate = y,
    hours = x,
    intercept
  )

@@

```{r}
#| label: fig-bayes-int-sd
#| fig-cap: Bayesian censored normal model vs LMM (dashed) fit to subset of data
#| fig-height: 3
#| fig-width: 8
ggplot(
  res,
  aes(
    hours, estimate,
    group = intercept,
    color = intercept
  )
) +
  geom_line() +
  geom_line(
    data = res_lmm,
    linetype = "dashed"
  ) +
  geom_ribbon(
    aes(
      x = hours,
      y = NULL,
      ymin = lwr,
      ymax = upr,
      color = NULL
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) + 
  color_scale
```

Lastly, we'll also fit the interaction model using a censored normal distribution, again with 1000 participants.

@@

We then compare the censored normal model with the LMM and recreate @fig-lmm-pre-interaction. As shown in @fig-cens-pre-interaction, there's almost no difference between the models. The LMMs CIs are slightly narrower near the intercept.
```{r}
#| label: fig-cens-pre-interaction
#| fig-cap: Censored normal vs LMM with a continuous by continuous interaction between within-session change and pretest mood.
#| fig-height: 3
#| fig-width: 8
get_emm_pre <- function(fit) {
  emmeans(
    fit,
    ~ hours + pre,
    # data = dat2,
    at = list(
      hours = seq(0, 3, length.out = 100),
      pre = c(
        pre_sum$mean - 2 * pre_sum$sd,
        pre_sum$mean - pre_sum$sd,
        pre_sum$mean,
        pre_sum$mean + pre_sum$sd
      )
    ),
    lmer.df = "asymp"
  )
}
pre_sum <- dat2 |>
  summarise(
    pre_mean = median(pre), .by = pid
  ) |>
  summarise(
    mean = mean(pre_mean), sd = sd(pre_mean)
  )
emm_cens <- get_emm_pre(fit3_b) |>
  as.data.frame() |>
  mutate(
    model = "cens"
  ) |>
  rename(
    lwr = lower.HPD,
    upr = upper.HPD
  )
emm_lmm <- get_emm_pre(fit3_small) |>
  as.data.frame() |>
  mutate(
    model = "LMM"
  ) |>
  rename(
    lwr = asymp.LCL,
    upr = asymp.UCL
  )

p_cens_pre_interaction <- bind_rows(emm_cens, emm_lmm) |>
  mutate(
    pre = factor(
      pre,
      levels = sort(unique(pre)),
      labels = c("-2 SD", "-1 SD", "Mean", "+1 SD")
    )
  ) |>
  {
    \(d)
    ggplot(
      data = d,
      aes(
        hours,
        emmean,
        group = interaction(pre, model),
        color = pre,
        linetype = model
      )
    ) +
      geom_line() +
      geom_ribbon(
        data = filter(d, model == "cens"),
        aes(
          ymin = lwr,
          ymax = upr,
          fill = pre,
          color = pre,
        ),
        color = NA,
        alpha = 0.25,
        show.legend = FALSE
      ) +
      geom_ribbon(
        data = filter(d, model == "LMM"),
        aes(
          ymin = lwr,
          ymax = upr,
          fill = pre,
          color = pre,
        ),
        fill = NA,
        alpha = 0.15,
        show.legend = FALSE
      ) +
      scale_color_brewer(palette = "Set1") +
      scale_fill_brewer(palette = "Set1")
  }()
p_cens_pre_interaction
```
